### 2024-06-21 presentation/presentation.tex
added presentation

\documentclass{beamer}

\usepackage{xcolor}

\newcommand*{\brk}{\\[10pt]}
\newcommand*{\doublebrk}{\\[10pt]}

\title{Approximation: Greedy and Local Search}

\author{Zixuan Fan}
\institute{Technische Universität München}

\date{July 2024}

\begin{document}
\begin{frame}
    \titlepage   
\end{frame}

\begin{frame}
    \frametitle{Agenda}
    \begin{enumerate}
        \item Introduction
        \item Scheduling Problems
            \begin{itemize}
                \item Scheduling on Single Machines
                \item Scheduling on Multiple Machines 
            \end{itemize}
        \item Graph Problems 
            \begin{itemize}
                \item K-center Problem 
                \item Travelling Salesman Problem
            \end{itemize}
        \item Conclusion
    \end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{Purpose of Approximation}
    \begin{itemize}
        \item<1-> Some problems not efficiently solvable: NP-hardness
        \item<2-> Can we find a flawed solution?
        \item<3-> How flawed/good is this solution?
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Quick Recap: Approximation Ratio}
    $Opt^*$: the optimal solution \\ 
    $Opt$: the suboptimal solution we compute \\[10pt]
    For minimization problem, we have
    \begin{align*}
        \alpha = \dfrac{|Opt|}{|Opt^*|}
    \end{align*}
\end{frame}

\begin{frame}
    \frametitle{Techniques in Approximation}
    \begin{enumerate}
        \item<1-> Randomization: MAXSAT
        \item<2-> Dynamic Programming: Knapsack, Bin Packing
        \item<3-> Linear \& Integer Programming: Primal-Dual, Semidefinite Program 
        \item<4-> \color{red} Greedy \& Local Search \color{black}
    \end{enumerate}
\end{frame}

\begin{frame}{Greedy and Local Search}
    \begin{itemize}
        \item<1-> Both strategies attempt to make the best decision
        \item<2-> Greedy algorithms forms a solution step by step 
        \item<3-> Local Search starts search from an arbitrary solution
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{K-Center Problem}
    \textbf{Problem Statement:} Given $n$ points in a space, 
    choose $k$ centers from them such that the sum of distances between 
    each point and its nearest neighbor is minimized. 
\end{frame}

\begin{frame}{K-Center Problem - Formal Definition}
    $d(p, q)$: distance between $p, q$ \\ 
    $d(p, S) := \min_{q \in S} d(p, q)$ \\
    \textbf{Input:} $S, k$ \\
    \textbf{Output:} $C \subseteq S$ s.t. $|C| = k$ and  $\sum_{p \in S} d(p, S)$ 
    is minimized \\[20pt]
    \pause 
    \textbf{What is the distance function? \\ Or what property does the distance function have?}
\end{frame}

\begin{frame}{Algorithm - Intuition}
    How do we search \textbf{greedily}? \\[10pt]
    If we have a temporary center $|C| < k$, how to build $|C'| = |C| + 1$?
\end{frame}

\begin{frame}{Algorithm - Description}
    Find $v \in S \backslash C$ s.t. $d(v, C)$ is maximized. \brk
    $C' = C \cup \{v\}$ \brk
    \pause
    What is the cost per iteration? \pause $O(|S| \cdot |S \backslash V|) = O(n^2)$
\end{frame}

\begin{frame}{Algorithm - Description}
    But, what is $C$ at the beginning of the iteration? 
    \pause 
    \begin{enumerate}
        \item The center of all points? $\implies$ Needs another $O(n^2)$ cost 
        \item Any arbitrary point $\implies$ Constant cost
    \end{enumerate}
    \pause
    Is the $O(n^2)$ cost worth it? We will discuss it later.
\end{frame}

\begin{frame}{Analysis}
    \textbf{Quadratic} cost
    \pause\brk
    \textbf{Approximation ratio}
\end{frame}

\end{document}